---
title: "Test link between subject-verb order and interrogative position"
output: pdf_document
---

# Load libraries and data

```{r warning=F, message=F}
library(lme4)
library(sjPlot)
```

```{r echo=F}

getMEText = function(r,ef, wald=NULL){
  
  AIC = r[2,]$AIC
  loglikDiff = signif(diff(r$logLik),2)
  chi = round(r$Chisq[2],2)
  df = r$`Chi Df`[2]
  p = signif(r$`Pr(>Chisq)`[2],2)
  
  wald.text = ""
  
  if(!is.null(wald)){
    est = signif(wald[1],2)
    stder = signif(wald[2],2)
    t = signif(wald[3],2)
    wptext = ""
    if(!is.na(wald[4])){
      wptext = paste(", Wald p =",signif(wald[4],2))
    }
    wald.text = paste("beta = ",est,", std.err = ",stder, ", Wald t = ",t,wptext,';')
  }
  
  begin = 'There was no significant'
  if(p <0.1){
    begin = "There was a marginal"
  }
  if(p < 0.05){
    begin = 'There was a significant'  
  }
  
  
  return(paste(begin,ef,"(",wald.text,"log likelihood difference =",
               loglikDiff,", df = ",df,", Chi Squared =", chi,", p = ",p,")."))
}

logit2per = function(X){
  return(exp(X)/(1+exp(X)))
}

```

```{r}
d = read.csv("wals-language.csv/language.csv")
```

# Make variables
```{r}
d$IP = d$X93A.Position.of.Interrogative.Phrases.in.Content.Questions

d$IP[d$IP %in% c("3 Mixed")] = NA
d$IP[d$IP %in% c("")] = NA
d$IP = as.character(d$IP)
d$IP[d$IP=="1 Initial interrogative phrase"] = "Initial"
d$IP[d$IP=="2 Not initial interrogative phrase"] = "Non-Initial"
d$IP = factor(d$IP)

d$WO = d$X81A.Order.of.Subject..Object.and.Verb

d$SV = d$X82A.Order.of.Subject.and.Verb
d$SV[d$SV==""] = NA
d$SV[d$SV=="3 No dominant order"] = NA
d$SV = d$SV == "1 SV"




d$Af = d$X26A.Prefixing.vs..Suffixing.in.Inflectional.Morphology
d$Af[d$Af %in% c("1 Little affixation","4 Equal prefixing and suffixing","")]=NA
d$Af = as.numeric(factor(d$Af))


```

Contingency table of initial interrogative phrase by basic word order:

```{r}
table(d$WO,d$IP) / rowSums(table(d$WO,d$IP))
```

Pattern:

-  Subject before verb  = non-initial
-  Verb before subject = initial


# Mixed effects model

Predict order of subject and verb by interrogative phrase position, controlling for family and macroarea. 

```{r}
m0 = glmer(IP ~ 1 + (1 + SV | family) + (1 + SV| macroarea), 
           data=d[!is.na(d$IP),], family = binomial)
m1 = glmer(IP ~ 1 + SV + (1 + SV| family) + (1 + SV| macroarea), 
           data=d[!is.na(d$IP),], family = binomial)
anova(m0,m1)
```

`r getMEText(anova(m0,m1), "main effect of subject-verb order", summary(m1)$coef['SVTRUE',])`.  Probability of initial interrogatives with subject-verb order = `r logit2per(fixef(m1)[1])`.  Probability of initial interrogatives with verb-subject = `r logit2per(sum(fixef(m1)))`

Look at direction:

```{r}
summary(m1)
```

## Random effects

Note that there are very small differences between random slope for SV by macroarea.  The convergence is singulative (total correlation between intercept and slopes), but there are no qualitative differences when removing the random slopes.


```{r fig.height=8}
sjp.glmer(m1, 're',sort.est = "(Intercept)", geom.colors = c(1,1))
```

